Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-11-16 manuscript/manuscript.tex
Specify arbitrary continuous density function

To take into account the curvature of the Earth, many global forward modelling methods
are defined in geocentric spherical coordinates.
A common approach is to discretize the Earth into tesseroids (spherical prisms),
which are defined by pairs of latitude, longitude, and
radial boundaries (see Fig.~\ref{fig:tesseroid}).
The gravitational fields generated by an arbitrary
tesseroid on any external point are given by volume
integrals that must be numerically approximated.
The literature offers two main approaches: one involves Taylor series expansion
\citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre
Quadrature (GLQ)
\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016}.
The Taylor series expansion is not well suited to develop an algorithm for
a density varying with depth according to an arbitrary function.
Different series expansion terms would have to obtained for each density function
desired.
Conversely, an arbitrary density function can be included in the GLQ without
any change to the integration method.

@@

In summary, the bibliography is offering different methodologies for computing
gravitational fields of homogeneous tesseroids or with certain specific density
functions in depth.
Although any density profile could be approximated by one of these specific density
functions, one should acknowledge the error that such approximation would propagate into
the gravitational field in order to guarantee certain degree of accuracy.

@@

We define an external computation point $P(r, \phi, \lambda)$ in a geocentric spherical
coordinate system at a radius $r$, geocentric latitude $\phi$, and longitude $\lambda$
where the gravitational fields are going to be calculated.
The first and second derivatives of the gravitational potential are taken with
respect to the local north-oriented Cartesian coordinate system of $P$
(Fig.~\ref{fig:tesseroid}).
\citet{Grombein2013} provide efficient formulations for the volume integrals of the
gravitational potential and its first and second derivatives of a tesseroid with
homogeneous density.
Here, we will assume that the tesseroid has a density varying with $r$ according to an
arbitrary function $\rho(r)$.
Thus, the integrals for the gravitational fields are slightly modified to

@@

The numerical integration of an arbitrary density function introduces
a new type of problem:
the integration error from using only a few nodes to discretize the density function.
The adaptive discretization may help
to reduce this kind of error.
However, it does not take into account the density function
and hence it is not well suited to fully perform this task.

>>>>

To take into account the curvature of the Earth, many global forward modelling methods
are defined in geocentric spherical coordinates.
A common approach is to discretize the Earth into tesseroids (spherical prisms),
which are defined by pairs of latitude, longitude, and
radial boundaries (see Fig.~\ref{fig:tesseroid}).
The gravitational fields generated by an arbitrary
tesseroid on any external point are given by volume
integrals that must be numerically approximated.
The literature offers two main approaches: one involves Taylor series expansion
\citep{Heck2007, Grombein2013} while the other makes use of Gauss-Legendre
Quadrature (GLQ)
\citep{Asgharzadeh2007, Wild-Pfeiffer2008, Li2011, Uieda2016}.
The Taylor series expansion is not well suited to develop an algorithm for
a density varying with depth according to an arbitrary continuous function.
Different series expansion terms would have to obtained for each density function
desired.
Conversely, an arbitrary continuous density function can be included in the GLQ without
any change to the integration method.

@@

In summary, the bibliography is offering different methodologies for computing
gravitational fields of homogeneous tesseroids or with certain specific density
functions in depth.
Although any continuous density profile could be approximated by one of these specific
density functions, one should acknowledge the error that such approximation would
propagate into the gravitational field in order to guarantee certain degree of accuracy.

@@

We define an external computation point $P(r, \phi, \lambda)$ in a geocentric spherical
coordinate system at a radius $r$, geocentric latitude $\phi$, and longitude $\lambda$
where the gravitational fields are going to be calculated.
The first and second derivatives of the gravitational potential are taken with
respect to the local north-oriented Cartesian coordinate system of $P$
(Fig.~\ref{fig:tesseroid}).
\citet{Grombein2013} provide efficient formulations for the volume integrals of the
gravitational potential and its first and second derivatives of a tesseroid with
homogeneous density.
Here, we will assume that the tesseroid has a density varying with $r$ according to an
arbitrary continuous function $\rho(r)$.
Thus, the integrals for the gravitational fields are slightly modified to

@@

The numerical integration of an arbitrary continuous density function introduces
a new type of problem:
the integration error from using only a few nodes to discretize the density function.
The adaptive discretization may help
to reduce this kind of error.
However, it does not take into account the density function
and hence it is not well suited to fully perform this task.
